---
title: "Learn History from 2010 China Population"
output: html_document
---
```{r, echo=FALSE}
library(knitr)
opts_chunk$set(message = FALSE, echo = FALSE, warning = FALSE)
```

The latest census in China was taken in 2010. The data are available online from
[National Bureau of Statistics of the China](http://www.stats.gov.cn/tjsj/pcsj/rkpc/6rp/indexch.htm). Although this
five-year-old census maybe too acient to be used for looking into the future of
the fast changing China, it is actually a reliable source to look back into China's 
history, especially when more detailed historical data are not accessible.  
    
In this post we will look at population data. The following
figure plots the number of people against the year of birth. Note that the numbers 
are of those who were still alive in 2010, not the total number of people born in 
that year. Actually there is no way to know the number of birth prior to 1949, when 
China was constantly at wars. 
```{r}
library(ggplot2)
source("global_functions.R")

# need to clean the excel file before reading
# remove unused rows and columns
# fill in missing cells, read_excel does not handel missing cells
population <- read_cencus_age("population_total_sect3_1.xls",
                              colNumber = 4, skip = 0)

# read_excel() does not read "0" into to numeric, quite annoying
population[3,1] <- 0

# make data ready for plot
population <- clean_cencus(population, 3, 4, c("age", "population"))
population$birthYear <- 2010 - population$age

# plot male and female side by side
g <- ggplot(data = population, aes(x = birthYear, y = population / 1e6))
g + geom_line(aes(color = sex)) + 
    geom_point(aes(color = sex)) +
    scale_x_continuous(breaks = seq(1900, 2020, by = 20) ) + 
    ggtitle("2010 pupulation") +
    xlab("Birth Year") +
    ylab("Population (million)")
```

The most striking feature in this figure is that the population change over birth
year is so violent. Normally, we would expect a smooth change over birth age, with 
fewer at older ages and more at younger ages, as a result of natural death and natural poplation growth. 

The violent changes in population indicates dramatical events. The first events 
occured around 1960, where there is a sudden drop in population. This event is 
officially called three-year natural disaster , or three-year difficult period, 
by the government, or called three-year great famine by many. There are still fier??ce
debate over how many people were starved to death as official numbers are not 
available, this plot clearly shows that up to 40% fewer people survived to 2010 born 
in those years. This could be due to much fewer baby were born or the infant fatality
is way too high. The population recovered quickly in 1962, overshot in 1963, and back
to normal trend until 1970.  
  
So what happend after 1970? 


